---
title: "Homework 3"
author: "Abigail Bolterman"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(alr4)
library(dplyr)
```

# Question 4.1
```{r}
bgsgirls_altered = BGSgirls %>%
  mutate(ave = (WT2+WT9+WT18)/3, lin = WT18-WT2, quad = WT2-2*WT9+WT18)
model = lm(BMI18 ~ ave + lin + quad, data = bgsgirls_altered)

summary(model)
```

The intercept is pretty much the same, with the results in 4.1 having an intercept of 8.298 versus the intercept of the transformed model of 8.310. 

The only coefficient besides the intercept with much significance is `lin` with a coefficient of 0.337. `ave` and `quad` are both less significant and are both negative and some similarities to results in 4.1. `ave` is similar to `WT2` in Model 2. `quad` is close to `WT18` in Model 1 and 3 and `DW18` in 2.


# Question 4.2
```{r}
# Set Up & Summaries
transact_altered = Transact %>%
  mutate(a = (t1+t2)/2, d = t1-t2)

m1 = lm(time ~ t1+t2, data = transact_altered)
m2 = lm(time ~ a + d, data = transact_altered)
m3 = lm(time ~ t2 + d, data = transact_altered)
m4 = lm(time ~ t1 + t2 + a + d, data = transact_altered)

summary(m1)
summary(m2)
summary(m3)
summary(m4)
```


## 4.2.1

Because `a` and `d` are linear combinations of `t1` and `t2`, the effect they would have on the model are already factored in with the inclusion of `t1` and `t2`. So their coefficients are labeled as `NA`.

## 4.2.2

The intercept for all four regressions are the same, as well as R-square, residual standard error, and F-statistic. The  estimations for `t1` and `t2` are the same in `m1` and `m4`. The estimates for `a` in `m2` and `t2` in `m3` are the same. The estimate for `d` in `m3` is the same as `t1` in `m1` and `m4`.


## 4.2.3

The difference is the use of `d` in place of `t1`. This makes a difference because `t2` is a component of `d`, so there's interaction between the two. 


# Question 4.5

Using $log_{10}(Y)$ changes the interpretation of a regression coefficient as a percentage change because what we've been referring to as $log(Y)$, which in other uses would represent $log_{10}(Y)$, actually means $log_{e}(Y)$, or $ln(Y)$. Because of this, the relationship between the variables and the response will be different.



# Question 4.6

The coefficient of `pctUrban` ($-0.01$) tells us that if the percent of people living in urban environments increases by 1, then the `fertility` will decrease by $e^{(0.01)} = 1.01$.


# Question 4.7

```{r}
model47 = lm(log(fertility) ~ log(ppgdp) + lifeExpF, data = UN11)
coefficients(model47)
```

With the fitted model, we get an estimated equation of 

$log(fertility) = 3.50736 - 0.06544*log(ppgdp) -0.02824*lifeExpF$.

To verify that a 25% increase in `ppgdp` will result in a 1.4% decrease in `fertility`, we will use the mean of `ppgdp` as our base value, increase it by 25%, then compare the resulting fertility values using the mean value for our `lifeExpF` variable.

```{r}
ppgdp_mean = mean(UN11$ppgdp)
ppgdp_mean_increased = 1.25 * ppgdp_mean
lifeExpF_mean = mean(UN11$lifeExpF)
fertility_base = exp(3.50736 - 0.06544*log(ppgdp_mean) -0.02824*lifeExpF_mean)
fertility_ppgdp_increased = exp(3.50736 - 0.06544*log(ppgdp_mean_increased) -0.02824*lifeExpF_mean)

((fertility_ppgdp_increased-fertility_base)/fertility_base)*100
```

Thus, we get a change of -1.4%.


# Question 4.8

Using equation 4.8 from the book, we have 
$3 + 4x_1+2E(X_2|X_1=x_1) < 0$

$\frac{3 + 4x_1+2E(X_2|X_1=x_1)}{2} < \frac{0}{2}$

$\frac{3}{2} + 2x_1 + E(X_2|X_1=x_1) < 0$

$E(Y|X_1=x_1) = -\frac{3}{2} - 2x_1 - E(X_2|X_1=x_1)$

# Question 4.9
## 4.9.1
The estimated salary for a male faculty member is $24,697 while the estimated salary for a female faculty member is 3340 less, resulting in a salary of 
$21,357.

## 4.9.2

This could happen if the effect of `Years` on female faculty is less than the effect on males, which would cause a less steep slope and move the intercept, in this case moved to be 201 units above the males' intercept.  












