---
title: 'Stat 305 Lecture 1 Part 2 Exercises: Matrices, Lists, and Data Frames, by Thomas Fiore'
author: "Solutions by Abigail Bolterman"
date: "2022/09/08"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Matrices

Create a $3\times 2$ matrix `A` with first column 1,2,3 and second column 4,5,6. 
```{r}
A = matrix(1:6, nrow=3)
A
```

Create a "column vector" `v` with entries 7,8. Hint: You can't code this as a vector, it is a $2 \times 1$ matrix `v`.
```{r}
v = matrix(7:8, ncol=1)
v
```


Then matrix multiply in R to get `Av` and compare that with your computation by hand. 
```{r}
R = A %*% v
R
```

Find the inverse of the $2 \times 2$ matrix with first row 1,2 and second row 3,4. And multiply `B` and `Binverse` in both orders to confirm that you get the identity. 
```{r}
B = matrix(1:4, nrow=2)
B
B_inverse = solve(B)
B_inverse
B%*%B_inverse
```

## Lists

Create a list `student` with `firstname` Joe, `entrance_year` 2018, and `honors` true. 
```{r}
student = list(firstname='Joe',
               entrance_year=2018,
               honors=TRUE)
student
```

Access the honors status in 3 ways as a logical.  
```{r}
student$honors
student[['honors']]
student[[3]]
```

Access the honors status in 2 ways as a sublist.  
```{r}
student['honors']
student[3]
```

## Data Frames


Try out the useful data frame commands on the `iris` data set. These commands were `?`, `View`, `names`, `head`, `tail`, `str`, `summary`, `pairs`, and `boxplot`. 

```{r}
View(iris)
names(iris)
head(iris)
tail(iris)
str(iris)
summary(iris)
pairs(iris)
boxplot(iris)
```

Extract the column of petal widths as a vector in 5 different ways. 
```{r}
iris$Petal.Width
iris[['Petal.Width']]
iris[[4]]
iris[,'Petal.Width']
iris[,4]
```


Extract entries 5 through 10 of the petal widths column as a data frame in two ways. 

```{r}
iris[5:10, 'Petal.Width']
iris[5:10, 4]
```

Extract the sub data frame of just setosa measurements. 

```{r}
iris[iris$Species == 'setosa', ]
```

Create a new data frame `iris_sepalwidth01` that has an additional column called `sepalwidth01` with 0 or 1 that indicates if the sepal width is above average. 

```{r}
iris_sepalwidth01 = iris
iris_sepalwidth01$sepalwidth01 = numeric(length(iris$Sepal.Width))
iris_sepalwidth01[iris$Sepal.Width > mean(iris$Sepal.Width), 'sepalwidth01'] = 1
iris_sepalwidth01
```
