---
title: "Stat 305 Lecture 13 MORE Exercises: Control Statements and Loops"
author: "Solutions by Abigail Bolterman"
date:   "   "
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## External Requirements

```{r warning=FALSE, message=FALSE}
# no packages required for control statements and loops
```

## Exercises

**1) Implement a fizzbuzz function. It takes a single number as input. if the number is divisible by three, it returns "fizz". If the number is divisible by five it returns "buzz". If the number is divisible by three and five, it returns "fizzbuzz". Otherwise, it returns the number. Make sure your function will notify the user in case a missing value is supplied or a vector of more than 1 values is supplied.**


```{r}
fizzbuzz <- function(x) {  
  if (length(x) >1){print("That's a vector.")}
  else if (is.na(x)) {print("Your input was an NA")}
  else if (x %% 3 == 0 & x %% 5 == 0){return("fizzbuzz")}
  else if (x %% 3 == 0){return("fizz")}
  else if (x %% 5 == 0){return("buzz")}
  else {return (x)}
}


fizzbuzz(c(1,2))
fizzbuzz(NA)
fizzbuzz(15)
fizzbuzz(6)
fizzbuzz(10)
fizzbuzz(7)
```

Make sure to test your function

**2) What happens if you use switch() with numeric values? Rewrite the function below with switch to figure it out** 

```{r}
print.rating <- function(service)
if(length(service)==1){
  if(service==1){
  print("poor")
} else if(service ==2){
  print("neutral")
} else if(service ==3){
  print("excellent")
} else{
  print("Invlaid choice, please enter either 1,2 or 3")
}
} else{
  print("service is supposed to be of length 1")
}
```


**3) What does the switch() call do? what happens if x is "e"**

```{r}
#switch(x,
#       a= ,
#       b= "ab",
#       c= ,
#       d = "cd")
```



